Best practices for forecasting changes in biodiversity: an example using breeding birds

### 2017-08-24 manuscript/draft.Rmd
remove duplicate ref, tidy others
minor edit
this didn't make sense

**Past environmental data.** Environmental data included a combination of
elevation, bioclimatic variables and a remotely sensed vegetation index (the
normalized difference vegetation index; NDVI), all of which are known to
influence richness and distribution in the BBS data [@kent2014]. For each year
in the dataset, we used the 4 km resolution PRISM data [@prism] to calculate
eight bioclimatic variables identified as relevant to bird distributions
[@harris2015]: mean diurnal range, isothermality, max temperature of the warmest
month, mean temperature of the wettest quarter, mean temperature of the driest
quarter, precipitation seasonality, precipitation of the wettest quarter, and
precipitation of the warmest quarter. Satellite-derived NDVI, a primary
correlate of richness in BBS data [@hurlbert2002], was obtained from the NDIV3g
dataset with an 8 km resolution [@pinzon2014] and was available from 1981-2013.
Average summer (May, June, July) and winter (December, January, Feburary) NDVI
values were used as predictors. Elevation was from the SRTM 90m elevation
dataset [@jarvis2008] obtained using the R package raster [@hijmans2016].
Because BBS routes are 40-km transects rather than point counts, we used the
average value of each environmental variable within a 40 km radius of each BBS
route's starting point.

@@

**Multi-timescale forecasts.** Another challenge for ecological forecasting is
the timescales of typical forecasts. We evaluated forecasts using yearly
timesteps up to a decade into the future. In contrast, many ecological forecasts
are temporally aggregated to 5-30 year timesteps and projected up to a century
into the future. These are commonly beyond the career or even lifespan of the
researchers, which makes it impossible to directly validate the models'
predictions. Currently, assessments at these long time-scales can only be made
with a small number of opportune datasets. For example, hindcasts have been used
to assess models of species richness [@algar2009; @distler2015] and distribution
[@rapacciuolo2012; @moran-ordonez2017; @araujo2005; @eskildsen2013] using data
aggregated across similar time scales to those studied here. More evaluation
studies like these are needed but there is a paucity of adequate long term data
available. Short term forecasts will allow for relatively rapid validation of
models by researchers without needing to wait for enough data to accumulate
[@tredennick2016; @dietze2016]. This does not mean that they should replace
large scale long-term forecasts. As discussed above, drivers of species richness
are likely to differ at different temporal scales [@rosenzweig1995; @white2004;
@white2007]. Thus, short-term forecasts should be expected to add value to
scientists and policymakers, and inform (not replace) long-term forecasts.
Hopefully, by evaluating trends in the shorter term forecasts as done here it
will be possible to also use the information from short-term forecasting to
improve longer-term estimates.

>>>>

**Past environmental data.** Environmental data included a combination of
elevation, bioclimatic variables and a remotely sensed vegetation index (the
normalized difference vegetation index; NDVI), all of which are known to
influence richness and distribution in the BBS data [@kent2014]. For each year
in the dataset, we used the 4 km resolution PRISM data [@prism] to calculate
eight bioclimatic variables identified as relevant to bird distributions
[@harris2015]: mean diurnal range, isothermality, max temperature of the warmest
month, mean temperature of the wettest quarter, mean temperature of the driest
quarter, precipitation seasonality, precipitation of the wettest quarter, and
precipitation of the warmest quarter. Satellite-derived NDVI, a primary
correlate of richness in BBS data [@hurlbert2002], was obtained from the NDIV3g
dataset with an 8 km resolution [@pinzon2014] and was available from 1981-2013.
Average summer (May, June, July) and winter (December, January, Feburary) NDVI
values were used as predictors. Elevation was from the SRTM 90m elevation
dataset [@jarvis2008] obtained using the R package raster [@raster].
Because BBS routes are 40-km transects rather than point counts, we used the
average value of each environmental variable within a 40 km radius of each BBS
route's starting point.

@@

**Multi-timescale forecasts.** Another challenge for ecological forecasting is
the timescales of typical forecasts. We evaluated forecasts using yearly
timesteps up to a decade into the future. In contrast, many ecological forecasts
are temporally aggregated to 5-30 year timesteps and projected up to a century
into the future. These are commonly beyond the career or even lifespan of the
researchers, which makes it impossible to directly validate the models'
predictions. Currently, assessments at these long time-scales can only be made
with a small number of opportune datasets. For example, hindcasts have been used
to assess models of species richness [@algar2009; @distler2015] and distribution
[@rapacciuolo2012; @moran-ordonez2017; @araujo2005; @eskildsen2013] using data
aggregated across similar time scales. More evaluation
studies like these are needed but there is a paucity of adequate long term data
available. Short term forecasts will allow for relatively rapid validation of
models by researchers without needing to wait for enough data to accumulate
[@tredennick2016; @dietze2016]. This does not mean that they should replace
large scale long-term forecasts. As discussed above, drivers of species richness
are likely to differ at different temporal scales [@rosenzweig1995; @white2004;
@white2007]. Thus, short-term forecasts should be expected to add value to
scientists and policymakers, and inform (not replace) long-term forecasts.
Hopefully, by evaluating trends in the shorter term forecasts as done here it
will be possible to also use the information from short-term forecasting to
improve longer-term estimates.
